---
title: '"Getting tuftesque with blogging"'
author: Shamindra Shrotriya
date: '2017-08-09'
slug: getting-tuftesque-with-blogging
categories:
  - data-science
  - stats
tags: []
---

```{r setup, echo=FALSE, cache=TRUE}
library(tufte)
```

## The tuftesque blogdown theme

[Nick Strayer](http://nickstrayer.me) and [Lucy D'Agostino McGowan](http://lucymcgowan.com) created the awesome [tuftesque blogdown theme](https://github.com/LFOD/real-blog/tree/master/themes/tuftesque) upon 
which this blog is based. I decided to finally bite the bullet and start
my blog with this wonderful theme. See [credits]() for additional sources of 
inspiration.

## Going 'full tufte' with blogging

Nick and Lucy already wrote a [blogpost already](http://livefreeordichotomize.com/2017/07/13/introducing-the-tuftesque-blogdown-theme/) detailing the process to get started with `blogdown` and 
the `tuftesque` theme. Please check it out as it is much likely 
to be more thorough than the instructions I've listed below.

I've added a bit more detail on how to launch your blog using [netlify](https://app.netlify.com/) and also setting up HTTPS and buying your
own domain from Google domains (and syncing with netlify). I've repeated
some of the common blogdown instructions to keep it all together.

## Instructions - your own tuftesque blog

1. You will need to install pandoc. Follow the official [installation instructions](http://pandoc.org/installing.html)

On Mac I just ran the following in the terminal:
```{r, eval=FALSE}
brew install pandoc
```

2. Open up Rstudio and install the [blogdown package](https://github.com/rstudio/blogdown) by running:

```{r, eval=FALSE}
devtools::install_github("rstudio/blogdown")
library(blogdown)
```

3. Install [hugo](https://gohugo.io/) by running:

```{r, eval=FALSE}
blogdown::install_hugo()
```

4. Initialize a new blog using blogdown by running:

```{r, eval=FALSE}
blogdown::new_site()
```

You should now see a new basic website appear in the RStudion viewer pane - yay
for us!

Now you can hit cancel in the Rstudio console pane `r emo::ji("stop_sign")`. 
This will stop the current site from being served (i.e. rendered live) in the 
html viewer pane in RStudio. I know it feels lame to shut it off so soon, 
but we are just getting started here!

5. Load the `tufteesque` theme. This is the moment you have all been waiting for
`r emo::ji("rainbow")`

```{r, eval=FALSE}
blogdown::install_theme('nstrayer/tuftesque')
```

6. Re-load the site to check the theme is updated:

```{r, eval=FALSE}
blogdown::serve_site()
```

7. Build site including all `Rmd` files - use method = "html"

When you 
This will knit the `Rmd` files into `html`
# See here for more details: https://bookdown.org/yihui/blogdown/methods.html
blogdown::build_site(local = FALSE, method = c("html"))

8. Create the gitignore file so that we commit only the relevant 
files for blogdown purposes. 

In the terminal run the following to create the `gitignore` file

```{r, eval=FALSE}
touch .gitignore
```

Inside the `gitignore` file using your favorite editor (i.e. `emacs` of course)
you can paste the following:

```{r, eval=FALSE}
# Blogdown common gitignore files
# Source: https://github.com/rbind/yihui/blob/master/.gitignore

.Rproj.user
.Rhistory
.RData
.Ruserdata
public
static/figures
blogdown

# Source: https://github.com/kbroman/blog/blob/source/.gitignore

_cache/
drafts/


# DS_Store files

public/.DS_Store
public/css/.DS_Store
public/images/.DS_Store
themes/tuftesque/.DS_Store
themes/tuftesque/data/.DS_Store
themes/tuftesque/exampleSite/.DS_Store
themes/tuftesque/layouts/.DS_Store
themes/tuftesque/static/.DS_Store
themes/tuftesque/static/css/.DS_Store
themes/tuftesque/static/images/.DS_Store
```

Inside the gitignore you should paste the following
See [here](https://github.com/shamindras/ss_personal_blog/blob/master/.gitignore) for the most up to date version of this file.

If there is a more elegant way to recursively remove the mac `DS_Store` files 
from the `gitignore` please leave a comment below `r emo::ji("smile")`.

9. Build your awesome website!

Build site - use method = "html" in the `build_site` function. This will 
knit the `Rmd` files into `html` which does not happen by default.

```{r, eval=FALSE}
blogdown::build_site(local = FALSE, method = c("html"))
```

You can also run `blogdown::serve_site()` to view the rendered site in the 
RStudio viewer pane. Now you should see the 3 default posts rendered tuftestyle
in your blog.

10. Add and commit all of the git files and push to github

```{r, eval=FALSE}
git add -A
git commit -m "FIX: Create first (default tufte-esque) blog, with gitignore"
git push origin master
```


11. _Optional:_ Install the `tufte` package

This is useful later when you start writing awesome blog posts and writing those
sweet margin notes `r emo::ji("sunglasses")` like a boss!

`r tufte::margin_note(stringr::str_c(emo::ji("v"), "Some cool marginalia here", sep = " "))`

```{r, eval=FALSE}

install.packages("tufte")
```

## Further Resources

- [Yihui et. al's Blogdown Book](https://bookdown.org/yihui/blogdown/). This is
the definitive (and [frequently updated](TODO: insert link here with latest github commits)) resource on all things blogdown
- [Nick and Lucy's LFOD blog](http://livefreeordichotomize.com/2017/07/13/introducing-the-tuftesque-blogdown-theme/) which got me started down this tuftesque route!
- 
